---
title: 'Math&146 Lab #8'
author: "N. Paterno"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

library(tidyverse)
library(palmerpenguins)
```

For this lab we will be building a scatter plot and least squares regression line, build a linear regression model and make some predictions. 

We'll be using the *penguins* data set from the *palmerpenguins* package. As always, lets glimpse the data. 

```{r}
glimpse(penguins)
```

For the examples, I'll focus on the body_mass_g and flipper_length_mm variables. 

## Basic Plot

First, we'll remove missing values, focus on our variables and do a basic scatter to visually determine if there is a linear relationship. 

```{r}
plot_data <- penguins %>% 
  select(flipper_length_mm, body_mass_g) %>% 
  na.omit()

ggplot(plot_data, 
       aes(flipper_length_mm, body_mass_g))+
  geom_point()
```

While its not exact, there does appear to be one.

## Correlation Coefficient

We can confirm what the plot shows us by calculating the correlation coefficient. 

```{r}
cor(plot_data$body_mass_g,plot_data$flipper_length_mm)
```
For the data, we have $r=0.87$. This is a moderate-strong linear relationship.

## Overlaid Plot

Now we'll recreate our scatter plot with the least squares regression line overlaid on top of the data.

```{r}
ggplot(plot_data, 
       aes(flipper_length_mm, body_mass_g))+
  geom_point()+
  geom_smooth(method = "lm",
              se = FALSE)
```

## Build a Model and Make a Prediction

We're ready to build our model and predict. First, we'll use the *lm* function to build the linear model. 

```{r}
mass_mod <- lm(body_mass_g ~ flipper_length_mm, data = plot_data) 
```

To make our predictions, we'll need to write our own function that will evaluate a data point on our model. (This code will be needed for the assignment.)

```{r}
pred <- function(model, value){
  return(model$coefficients[1]+value*model$coefficients[2])
}
```

Finally, we'll predict the mass of a penguin with a flipper length of 200mm.

```{r}
pred(mass_mod, 200)
```

According to our model, a penguin with a flipper length of 200mm will have a body mass of 4156.282g.

## Assignment

In a script file: 

1. Clean the data by focusing on bill length and body mass then removing missing values.

2. Repeat the above process (basic scatter, correlation, overlaid scatter, model, prediction) to try and predict body mass from bill length. 

Save your work in a script file called: LastName_lab8. For example, my assignment file would be Paterno_lab8.R