---
title: 'Math&146 Lab #5'
author: "N. Paterno"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

library(tidyverse)
library(openintro)
```

This lab will be brief compared to our last few. We will look at a few different ways to plot our data to determine its shape as either normal, skewed left or skewed right. 

We'll be using the *speed_gender_height* data set from the *openintro* package. Let's see the data. 

```{r}
glimpse(speed_gender_height)
```

Here we have a data frame that includes a person's biological gender, height and the maximum speed they have personally driven a car. For the examples I will focus on the *height* variable. 

## Cleaning the Data

To clean the data, I'll remove observations where a height wasn't reported. 

```{r}
plot_data <- speed_gender_height %>% 
  na.omit()
```

## Visuallizing Shape

We'll look at three different plots that can help us visually determine the shape of our data: histogram, frequency polygon and a density curve. 

### Histogram

```{r}
ggplot(plot_data, 
       aes(height))+
  geom_histogram()
```

It appears that our data is approximately normal (bell shaped). 

### Frequency Polygon

```{r}
ggplot(plot_data, 
       aes(height))+
  geom_freqpoly()
```

While the curve is not smooth, it does show that the data is approximately normal. 

### Density Curve

```{r}
ggplot(plot_data, 
       aes(height))+
  geom_density()
```

Here, we essentially have a smoothed version of the frequency polygon. The major difference is that our y-axis is no longer a frequency but a percentage. 

## Assignment
In a script file: 

1. Clean the data to remove any observations that did not have a reported speed. 

2. Create a histogram, frequency polygon and density curve for speed using the cleaned data.

Save your work in a script file called: LastName_lab5. For example, my assignment file would be Paterno_lab5.R

## Extra Material
As with the last lab, we'll go a bit further to see if shape is independent of gender. 

We can do this using a *facet_wrap* which will create a separate plot for each gender. 

```{r}
ggplot(plot_data, 
       aes(height))+
  geom_density()+
  facet_wrap(~gender)
```

It appears that the data is still approximately normal when we focus on individual genders. 

Another way we could do this is using color. 

```{r}
ggplot(plot_data, 
       aes(height, 
           fill = gender))+
  geom_density(alpha = 0.5)
```

The *alpha* argument specifies how see-through the plot is. Closer to 1 is opaque, closer to 0 is transparent.