---
title: 'Math&146 Lab #2'
author: "N. Paterno"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

library(tidyverse)
```

For this lab, we will be looking at variable, types, arithmetic, data frames and a *little bit* of data wrangling. 

## Variables and Types

First we will look at variables. As we've seen, there are two classes of variables: numeric and categorical. In *R* there are many types of variables. Numeric variables are typically stored as *INTEGER* (discrete) or *DOUBLE* (continuous). Categorical variables are typically stored as *CHARACTER* or *FACTOR*.

In the code chunk below, we'll define a numeric variable and a categorical variable. 

```{r}
x <- 12
y <- "yellow"
```

The <- symbol is the assignment operator in *R*. There you can type < and - separately or use keyboard shortcuts. On Windows its alt- and a Mac its option-. 

Let's check how these are stored.

```{r}
typeof(x)
typeof(y)
```

The numeric variable, x, is stored as a double and the categorical variable, y, is a character. It is important to note that character variables should always use quotes. 

Since every data set we see has more than one entry, it would be useful to be able to store more than one value. Luckily, we can! These are called *vectors*.

```{r}
x_2 <- c(12, 10, 8, 6, 4, 2)
```

This store the *c*ollection of doubles in the variable x_2.

## Arithmetic

We saw in lab #1 that we can use *R* like a giant, overpowered calculator. We can do this with variables too. 

```{r}
x+5
x/24
```

We can do arithmetic with vectors too!

```{r}
x_2+5
x_2/24
```

## Data Frames

When we work with data sets in *R*, we are viewing them as data frame. A data frame is a structure where each column is a variable and each row is an observation.

Here is a data frame of my immediate family. First I will need to define each variable as a vector and then join them together into a data frame. The *head* function will show us the first six observations in the data frame.

```{r}
name <- c("Jim", "Peggy", "Kyla", "Nick", "Alex", "Tori", "Belinda", "Nathan", "Kelsey", "Tommy")

relationship <- c("dad", "mom", "sister", "me", "brother", "sister", "wife", "son", "sister-in-law", "brother-in-law")

age <- c(58, 58, 36, 33, 31, 29, 34, 2, 32, 29)

family <- data.frame(name, age, relationship)

head(family)
```

## Data Wrangling

Data wrangling is the process of 'cleaning' or manipulating a data set into a form that we can analyze. For this section, we will use the *gov_poll* data set from the *openintro* package. Let's look at the data set first. 

```{r}
example_data <- openintro::gov_poll

head(example_data)
```

This appears to be an approval poll of President Obama. Let's confirm this. We'll use the *unique* function which will give us all unique values of a variable. The input for this function the dataframe$variable.

```{r}
unique(example_data$poll)
unique(example_data$eval)
```

There was more than we thought! This data set *is* an approval poll. However, it is not just for President Obama; it also includes Democrats and Republicans (presumably those in congress).

For our intro to wrangling, we will create a new data frame that focuses on Obama and then calculate his approval/disapproval percentages. 

```{r}
obama <- example_data %>% 
  filter(eval == "Obama")
```

There are a few new things in the above code. First, the pipe operator: %>% . This can be read as "and then" so our new *obama* data frame will start with the *example_data* "and then" will be filtered to only include his approval and disapproval votes. 

The *filter* function input is a logical equation (hence the double =). It will look at each row and determine if it is true or false. True observations are included and false are left out. 

Next, we'll calculate the percentages. To do so we'll need to know how many votes there are total and how many are approval and disapproval. We'll use *group_by* to isolate approve/disapprove, then *summarize* to count the votes for approve/disapprove and finally *mutate* to create new variables for the total number of votes and percentages. 

```{r}
obama %>% 
  group_by(poll) %>% 
  summarize(count = n()) %>% 
  mutate(total = sum(count), 
         percent = count/total)
```

This shows us that, at the time, the President had a 57.75% approval rating and 42.25% disapproval rating. 

## Assignment
In a script file: 

1. Create a data frame consisting of three people that includes name and age. 

2. Calculate the approval/disapproval ratings for Democrats or Republicans. (This problem is *almost* a copy/paste.)

Save your work in a script file called: LastName_lab2. For example, my assignment file would be Paterno_lab2.R